\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model

### 2017-10-29 ASApaper/paper-child.Rnw
mini changes

We want to explore three different aspects of the SAOM models using the lineup protocol: significance of parameters, goodness-of-fit of a model, and visual detection of additional parameters. Each on of these requires a different setup, which we describe in detail, but we use the lineup protocol to examine each aspect. In each lineup, there are two models shown: the null model and the alternative model. The definition of the null and alternative model will vary according to which aspects of the SAOMs we are exploring. 
%For each lineup, we consider two models: model M1 is the null model, and another model, from M2-M5, which is chosen to be the alternative model.

To construct a lineup, we simulate five networks from the null model and one network from the alternative model. We chose to expose our participants to only six plots at a time in order to show the node-link diagrams in more detail and to lower their cognitive load. Typically, lineup experiments are done with sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 node-link diagrams. An example of a lineup like those shown to our participants is shown on the right side of the image  in Figure~\ref{fig:shiny}.  In this lineup, model M4 is the alternative model, and model M1 is the null model. To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$. For $\beta_5$, twice the estimated value was used. More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvd}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}. To get the simulations, we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.

\begin{figure}
\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
\caption{\label{fig:shiny} A screen grab of the web application we created to design our lineup experiment. More details about this application are given in Section~\ref{sec:setupvd}. In the lineup, M4 is the alternative model and one estimate from model M4 with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues} is placed at random among five observations simulated from the null model, M1. Which plot comes from model M4?  %The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
}
\end{figure}

@@

\begin{description}
\item $H_0$: $\beta_k = 0 $
\item $H_A$: $\beta_k \neq $
\end{description}

@@

The lineup generated under this scenario is shown to independent viewers. If the viewers pick out the alternative data plot, that is evidence in favor of the null hypothesis, while picking one of the null plots is evidence against the null hypothesis. The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M2 and M3, respectively.

@@

To generate the null plots, we simulate five networks from model M\textit{i} using the corresponding parameters in Table~\ref{tab:fittedvalues}. We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place the true second wave data. We cannot show the data more than once to each participant, so we examine several different models in our Amazon Mechanical Turk experiment, each participant never seeing the true data wave twice. The models we chose for goodness-of-fit testing are M2, M3, M4, and M6.

@@

Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do. Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM. By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model. An effect becomes noticeable when a simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent when they simulations are placed in a lineup. We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.

@@

To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive. We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}. The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup. To determine the six levels of the effects we want to test, started with the estimates of the parameter at hand as determined by the procedure described in Section~\ref{sec:models}, and multiply it or changing its sign to determine what \textit{we} notice about the effect of the parameter in simulations from the changed models.

In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{shiny}. A screen grab of the app is shown in Figure~\ref{fig:shiny}. On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams. There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong. The plots that appear that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ given in Table~\ref{tab:fittedvalues}.

>>>>

We want to explore three different aspects of the SAOM models using the lineup protocol: (1) significance of parameters, (2) goodness-of-fit of a model, and (3) visual detection of parameters. 
Each one of these situations requires a different setup, which we describe in detail, but we make ue of the lineup protocol for all of these aspects.

In each lineup, we include plots from two models: the null model and an alternative model. The definition of the null and alternative model varies with the aspect of the SAOMs we are exploring. 
%For each lineup, we consider two models: model M1 is the null model, and another model, from M2-M5, which is chosen to be the alternative model.

 Typically, a lineup shows sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 node-link diagrams. We chose to expose our participants to only six plots at a time in order to show the node-link diagrams in more detail and to lower cognitive load for participants.
To construct a lineup, we simulate five networks from the null model and one network from the alternative model.  An example of a lineup like those shown to our participants is given on the right side of the image  in Figure~\ref{fig:shiny}.  In this lineup, model M4 is the alternative model, and model M1 is the null model.

To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$. For $\beta_5$, twice the estimated value was used. More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvd}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}. To get the simulations, we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.

\begin{figure}
\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
\caption{\label{fig:shiny} A screen shot of the web application we created to design our lineup experiment. More details about this application are given in Section~\ref{sec:setupvd}. In the lineup, M4 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}. One plot simulated from this model is placed at random among five observations simulated from the null model, M1. Participants of the study are asked to identify the most different plot.  %The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
}
\end{figure}

@@

\begin{description}
\item $H_0$: $\beta_k = 0 $
\item $H_A$: $\beta_k \neq 0$
\end{description}

@@

The lineup generated under this scenario is shown to a number of independent viewers. If an observer picks out the alternative data plot, that is evidence in favor of the null hypothesis, while picking one of the null plots is evidence against the null hypothesis. The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M2 and M3, respectively.

@@

To generate the null plots, we simulate five networks from model M\textit{i} using the corresponding parameters in Table~\ref{tab:fittedvalues}. We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place a node-link diagram of the true second wave data. We cannot show the data more than once to each participant, so we examine several different models in our Amazon Mechanical Turk experiment, each participant never seeing the true data wave twice. The models we chose for goodness-of-fit testing are M2, M3, M4, and M6.

@@

Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do. Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM. By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model. We define an effect to be noticeable, if viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect. We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.

@@

To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive ones. We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}. The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup. To determine the six levels of the effects we want to test, started with the estimates of the parameter at hand as determined by the procedure described in Section~\ref{sec:models}, and multiply it or changing its sign to determine what \textit{we} notice about the effect of the parameter in simulations from the changed models.

In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{shiny}. A screen shot of the app is shown in Figure~\ref{fig:shiny}. On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams. There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong. The plots that appear that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ given in Table~\ref{tab:fittedvalues}.
